Human tissue specific markers

### 2017-10-17 manuscript/manuscript.Rnw
update manuscript figures

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{siunitx}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage[macros=false,single=true,single-form=long]{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage[export]{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
\usepackage[markifdraft,raisemark=0.01\paperheight,draft]{gitinfo2}
% this package will stretch the font slightly if necessary.
%\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\usepackage[capitalize]{cleveref}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\setlength{\emergencystretch}{20em}
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}
\ExplSyntaxOn
\NewAcroCommand \gene {
  \group_begin:
    \keys_set:nn {acro} {short-format = \itshape}
    \acro_use:n{#1}
  \group_end:
}
\NewAcroCommand \protein { \acro_use:n {#1} }
\ExplSyntaxOff

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(dev="cairo_pdf",cache=TRUE, autodep=TRUE)
options(scipen = -1, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("xtable")
library("caret")
library("pROC")
library("grid")
library("scales")

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced",
                   "housekeeping_genes_superset"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")

@@

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,compact_tissue:=factor(compact_tissue_name(tissue),
                                                       levels=compact_tissue_name(interesting.tissues))
                               ]
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    if (!is.null(tissue.name)) {
        gene.levels <- 
            unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
    } else {
        gene.levels <- 
            unique(as.character(tissue.specific.genes[tissue %in% interesting.tissues
                                                      ][order(specificity),gene]))
    }
    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,
                      x=compact_tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="black",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
                   axis.ticks = element_blank(),
                   panel.grid=element_blank(),
                   panel.border=element_blank(),
                   panel.background=element_blank(),
                   plot.background=element_blank(),
                   plot.margin=unit(c(1,1,1,1),"mm"),
                   axis.text.y = element_text(size=10,face="italic"),
                   axis.text.x = element_text(size=9,angle = 30, hjust = 1, vjust=1))
           )
}
@

@@

<<tissue.specific.markers.table,results="hide",echo=FALSE>>=
tissue.specific.markers.table <- function(tissue,legend,min.gi=0) {
    this.tissue <- tissue
    temp <-
        tissue.specific.genes[tissue==this.tissue & gi.max >= min.gi,
                              list(Gene=link.gene.names(gene),
                                   "{Tissue specificity}"=specificity,
                                   "{Mean Expression}"=gi.max)]
    setorder(temp,-`{Mean Expression}`)
    print(xtable(temp,
                 align=c("r","c","S","S[table-format=1.0]"),
                 caption=legend,
                 label=paste0("tab:",tissue,"_specific_genes")),
          sanitize.text.function=genes.table.sanitize,
          include.rownames=FALSE,
          booktabs=TRUE,
          floating=TRUE,
          )
    rm(temp)
}

@@

<<placenta.specific.markers.table,results="asis",echo=FALSE>>=
tissue.specific.markers.table(tissue="placenta",
                              legend=
                                  paste0("Placenta specific genes with expression ",
                                         "$\\ge 100$ ordered by median expression in placenta."),
                              min.gi=100
                              )
@

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{siunitx}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage[macros=false,single=true,single-form=long]{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage[export]{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
% \usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
\usepackage[markifdraft,raisemark=0.01\paperheight,draft]{gitinfo2}
% this package will stretch the font slightly if necessary.
%\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\usepackage[capitalize]{cleveref}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
\oddsidemargin 0.0in 
\textwidth 6in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
\setlength{\emergencystretch}{20em}
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}
\ExplSyntaxOn
\NewAcroCommand \gene {
  \group_begin:
    \keys_set:nn {acro} {short-format = \itshape}
    \acro_use:n{#1}
  \group_end:
}
\NewAcroCommand \protein { \acro_use:n {#1} }
\ExplSyntaxOff

@@

<<load.libraries,echo=FALSE,results="hide",warning=FALSE,message=FALSE,error=FALSE,cache=FALSE>>=
opts_chunk$set(out.width="0.5\\textwidth",out.height="0.7\\textheight",out.extra="keepaspectratio")
opts_chunk$set(dev="CairoPDF",cache=TRUE, autodep=TRUE)
options(scipen = -1, digits = 2)
library("data.table")
library("ggplot2")
library("reshape2")
library("proto")
library("xtable")
library("caret")
library("pROC")
library("grid")
library("scales")

@@

<<load.data,echo=FALSE,results="hide",warning=FALSE,cache.extra=file.info(paste0("../tissue_specific_expression/",c("tissue_specific_markers","all_interesting_gene_reads_wide","combined_star_logs","svm_test_results","svm_test_results_reduced","housekeeping_genes_superset")))[,"mtime"]>>=
for (a in paste0("../tissue_specific_expression/",
                 c("tissue_specific_markers",
                   "all_interesting_gene_reads_wide",
                   "combined_star_logs",
                   "svm_test_results",
                   "svm_test_results_reduced",
                   "housekeeping_genes_superset"
                   ))) {
    load(a)
}
tissue.specific.genes <- data.table(tissue.specific.genes)
setkey(tissue.specific.genes,"gene")

@@

    tissue.specific.gene.reads <- 
        data.table(melt(interesting.gene.reads.wide[genes,interesting.tissues,with=FALSE],
                        id="gene_short_name",
                        variable.name="tissue",
                        value.name="FPKM"
                        ))
    tissue.specific.gene.reads[,compact_tissue:=factor(compact_tissue_name(tissue),
                                                       levels=compact_tissue_name(interesting.tissues))
                               ]
    tissue.specific.gene.reads[,`log_2(FPKM)`:=log2(FPKM+1)]
    if (!is.null(tissue.name)) {
        gene.levels <- 
            unique(as.character(tissue.specific.genes[tissue==tissue.name][order(specificity),gene]))
    } else {
        gene.levels <- 
            unique(as.character(tissue.specific.genes[tissue %in% interesting.tissues
                                                      ][order(specificity),gene]))
    }
    tissue.specific.gene.reads[,gene_short_name:=factor(gene_short_name,levels=gene.levels)]
    return(ggplot(tissue.specific.gene.reads,
                  aes(y=gene_short_name,
                      x=compact_tissue))
           + geom_tile(aes(fill=`log_2(FPKM)`))
           + scale_fill_gradient(limits=c(0,17),
                                 low="white",
                                 high="black",
                                 guide=guide_colourbar(title=expression(log[2](FPKM))))
           + scale_x_discrete("", expand = c(0, 0))
           + scale_y_discrete("", expand = c(0, 0))
           + theme_grey(base_size=14)
           + theme(legend.position = "top",
                   axis.ticks.y = element_blank(),
                   axis.ticks.x = element_blank(),
                   panel.grid=element_blank(),
                   panel.border=element_blank(),
                   panel.background=element_blank(),
                   plot.background=element_blank(),
                   plot.margin=unit(c(1,1,1,1),"mm"),
                   axis.text.y = element_text(size=10,face="italic"),
                   axis.text.x = element_text(size=9,angle = 40,
                                              hjust = 1, vjust=1,
                                              margin=unit(c(0,0,0,0),"cm")))
           )
}
@

@@

<<tissue.specific.markers.table,results="hide",echo=FALSE>>=
tissue.specific.markers.table <- function(tissue,legend,min.gi=0) {
    this.tissue <- tissue
    temp <-
        tissue.specific.genes[tissue==this.tissue & gi.max >= min.gi,
                              list(Gene=link.gene.names(gene),
                                   specificity,
                                   gi=gi.max)]
    setorder(temp,-gi)
    print(xtable(temp[,list(Gene,
                            "{\\multicolumn{1}{p{1.5cm}}{\\centering Tissue \\\\ specificity }}"=
                            specificity,
                            "{\\multicolumn{1}{p{1.5cm}}{\\centering Mean \\\\ Expression }}"=
                                gi)],
                 align=c("r","c","S[table.format=1.2]","S[table-format=5.0,round-mode=places,round-precision=0]"),
                 caption=legend,
                 label=paste0("tab:",tissue,"_specific_genes")),
          sanitize.text.function=genes.table.sanitize,
          include.rownames=FALSE,
          size="\\footnotesize",
          booktabs=TRUE,
          floating=TRUE,
          )
    rm(temp)
}

@@

<<placenta.specific.markers.table,results="asis",echo=FALSE>>=
tissue.specific.markers.table(tissue="placenta",
                              legend=
                                  paste0("Placenta specific genes ($\\tau_g \\ge 0.98$) ",
                                         "with expression ",
                                         "$\\ge 300$ ordered by mean expression in placenta."),
                              min.gi=300
                              )
@
